---
#title: "Peer Assesment 1"
##author: "Santiago Armida"
##date: "12 de junio de 2015"
###output: html_document
---

The variables included in this dataset are:
  steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
  date: The date on which the measurement was taken in YYYY-MM-DD format
  interval: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

#  Reproductible Research
# by Santiago Armida: June 14, 2015
## read data and setup envirnment
##  it will read the data in current directory
### loads needed packages
- \usepackage{sqldf}
```{r}
#Install and load libraries
#  library(plyr)
#Install can be comented (#) if the package is already down
# packages to solve question via SQL
#   install.packages("sqldf")
   library(sqldf)
## read data  Activity.csv in same directory
    
    acts <- read.csv("activity.csv")
```
Calculate the mean and total steps per day SQL
```{r, echo=TRUE}
stepdata <- sqldf("select date, avg(steps) as avgstep, sum(steps) as sumsteps 
                  from acts
                  where steps <> 'NA'
                  group by date")

#stepsINT <- ddply(acts, .(interval), summarize, mean=mean(steps, na.rm = TRUE))
```
Histogram of total STEPS

```{r, echo=TRUE}
hist(stepdata$sumsteps,main="Histogram of total steps by day",xlab = "Steps in interval",ylab = "Distribution of steps",col = "green")
```

What is the average daily activity pattern?
```{r simultation, echo=TRUE}
stepxday <- sqldf("select interval, avg(steps) as avgstep, sum(steps) as sumsteps
                  from acts
                  where steps <> 'NA'
                  group by interval")
stepmaxint <- sqldf("select interval, avgstep
                    from stepxday
                    where avgstep =(select max(avgstep)
                                      from stepxday)")
stepmaxint
plot(stepxday$interval,stepxday$avgstep,type="l",main="Steps per day",xlab = "Steps in interval 00:00 to 24:00",ylab = "Distribution of steps",col = "blue")
```
Imputing missing values
The average for the interval is assigned to the NA value
A UNION of both data frames is performed

```{r, echo=TRUE}
numNA <- sqldf("select count(*) NUM from acts where steps is NULL")
numNA
NAs <- sqldf("select a.date, a.interval, b.avgstep from acts a, stepxday b where a.steps is NULL and a.interval = b.interval") 
ALLsteps <- sqldf("select date, interval, steps from acts a where a.steps is NOT NULL UNION select * from NAs order by 1,2")
stepdataB <- sqldf("select date, avg(steps) as avgstep, sum(steps) as sumsteps from ALLsteps  group by date")
hist(stepdataB$sumsteps,main="Histogram of total steps by day",xlab = "Steps in interval",ylab = "Distribution of steps",col = "orange")
summary(stepdataB)
```
Weekday analysis
##Function day of week not working in SQL
```{r, echo=TRUE}
```{r, echo=TRUE}
weekday = weekdays(as.Date(ALLsteps$date))
head(weekday)
weekdata <- cbind(ALLsteps, weekday)
head(weekdata)
```
```
